\documentclass[compress]{beamer}
\useoutertheme[footline=authortitle,subsection=false]{miniframes}
\setbeamertemplate{navigation symbols}{}

\definecolor{vertmoyen}{HTML}{A9CB60}
\definecolor{vertclair}{HTML}{D3F192}
\definecolor{vertfonce}{HTML}{347003}
\definecolor{deletecol}{HTML}{3B8A13}
\definecolor{addcolor} {HTML}{9E0270}

\usecolortheme[named=vertfonce]{structure}

\setbeamercolor{section in head/foot}{fg=vertfonce,bg=vertmoyen}
\setbeamercolor{author in head/foot}{fg=white}
\setbeamercolor{titlelike}{bg=vertclair}
\setbeamertemplate{mini frame in other subsection}[default][0!white!100]
\setbeamertemplate{section in head/foot shaded}[default][0!white!100]

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{kpfonts}

\usepackage{amsmath}

\newcommand{\dual}[1]{\ensuremath{\bar{#1}}}

\newcommand{\set}[2]{\ensuremath{\left\{#1\mid#2\right\}}}

\newenvironment<>{subpart}[1]
{ \begin{block}#2{#1}
  \begin{itemize}
}{
  \end{itemize}
  \end{block}
}

\newcommand<>{\singleline}[1]{\begin{block}{\only#2#1}\end{block}}

\usepackage[vlined]{algorithm2e}

\usepackage{tikz}
\usetikzlibrary{positioning}

\newcommand{\basecharttitle}[1]{{\color{vertfonce}#1}\vspace{-0.8em}}
\makeatletter
\newcommand{\charttitle}{\@ifstar\charttitlestar\charttitleplain}
\makeatother
\newcommand{\charttitleplain}[1]{\hspace{2em}\basecharttitle{#1}}
\newcommand{\charttitlestar}[1]{\centering\basecharttitle{#1}}

\newcommand{\compC}{Compression ratio comparison}
\newcommand{\timeC}{Compression time comparison \scriptsize (in seconds)}
\newcommand{\axioC}{Axiom compression ratio comparison}

\newcommand{\asGoodRPILU}{For a combined algorithm to be always at least as
good as RPI.LU it has to compute root and units safe literals.}
\newcommand{\asGoodLURPI}{For a combined algorithm to be always at least as
good as LU.RPI it has to be able to lower units introduced by RPI.}

\newcommand{\chartscale}{1}

\tikzstyle{proof edge}=[thick,cap=round]
\tikzstyle{deleted edge}=[proof edge, dashed, color=deletecol]

\newcommand{\proofnode}[3][]{
  \node [anchor=mid, #1] (#2) {#3}
}

\newcommand{\rootnode}[1][]{
  \proofnode[#1]{root}{$\bot$}
}

\newcommand<>{\drawchildren}[3]{
  \draw#4 [proof edge] (#1) -- (#2);
  \draw#4 [proof edge] (#1) -- (#3)
}

\newcommand{\addchildren}[5]{
  \proofnode[above left  of=#1]{#2}{#3};
  \proofnode[above right of=#1]{#4}{#5}
}

\newcommand{\withchildren}[5]{
  \addchildren{#1}{#2}{#3}{#4}{#5};
  \drawchildren{#1}{#2}{#4}
}

\newcommand<>{\crossnode}[2][]{
  \draw#3 [color=deletecol,thick,cap=round,#1] (#2.mid) ++(10:0.3) -- ++(190:0.6);
}

\newcommand<>{\pivot}[3][]{
  \node#4 [anchor=base,color=vertfonce,#1] at (#2.north) {\scriptsize #3};
}

\title{Combining RecyclePivotsWithIntersection and LowerUnits}
\author[Joseph Boudou]{Joseph Boudou, Bruno Woltzenlogel Paleo}
\institute[GSoC]{Google Summer of Code 2012}
\date{October 2012}

%\includeonly{}
%%%%%%%%%%%%%%%%%%
% Begin Document %
%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Overview}
\tableofcontents
\end{frame}

\begin{frame}{Benchmarks}
  \singleline{Algorithms implemented in Scala for Skeptik}
  \begin{subpart}{1000 proofs from VeriT}
    \item 100 SAT proofs from the old external SAT solver
    \item 900 SMT proofs translated to propositional resolution
  \end{subpart}
\end{frame}

\section{Introduction}
\subsection{RPI, LU, composing them}

\begin{frame}{Propositional resolution calculus}
\begin{subpart}{Conventions}
  \item Clauses are sets of literals.
  \item Let $\eta_{\Gamma}$ be a proof of clause $\Gamma$ and $\eta_{\Delta}$ a proof of $\Delta$.
  \item $\eta_{\Sigma} = \eta_{\Gamma} \odot_a \eta_{\Delta}$ iff its clause $\Sigma$ is such that
        $\left\{\dual{a}\right\} = \Gamma \setminus \Sigma \wedge \left\{a\right\} = \Delta \setminus \Sigma$.
  \item $a$ is the \emph{pivot} of $\eta_\Sigma$.
  \item $\eta_\Gamma$ and $\eta_\Delta$ are the \emph{premises} of $\eta_\Sigma$.
  \item $\eta_\Sigma$ is a \emph{child} of both $\eta_\Gamma$ and $\eta_\Delta$.
\end{subpart}
\begin{subpart}{Proof as DAG}
  \item A node can have more than one child.
\end{subpart}
\end{frame}

\begin{frame}{Regular proof}
\begin{definition}[Regular proof]
  A proof is said to be regular if on every path from its root to any of its axiom, each pivot appears only once.
\end{definition}
\begin{theorem}<2->[Tseitin]
  Given a set of axioms and a clause $\Gamma$, the smallest regular proof of $\Gamma$ might be exponentialy bigger
  than the smallest irregular proof of $\Gamma$.
\end{theorem}
\end{frame}

\begin{frame}{Extending Irregularity}
\begin{definition}[Fully regular proof]
  A proof is fully regular if for each variable there is at most one resolution node with this variable as pivot.
\end{definition}
\begin{subpart}{Conventions}
  \item Usual irregularities are called \emph{vertical irregularities}.
  \item Other irregularities are called \emph{horizontal irregularities}.
\end{subpart}
\end{frame}

\begin{frame}{RecyclePivotsWithIntersection (RPI)}
  \begin{subpart}{Partial Regularization}
    \item Delete a branch only if the pivot appears on \textbf{every} path from the root to the node.
  \end{subpart}
  \begin{definition}[Safe literal]
    A literal is safe for a node $\eta$ if it can be added to $\eta$'s clause without changing the proof's conclusion.
  \end{definition}
  \begin{subpart}{Two traversals}
    \item Collect safe literals and mark edges to delete.
    \item Delete edges and fix the proof.
  \end{subpart}
\end{frame}

\begin{frame}{LowerUnits (LU)}
  \begin{subpart}{Lowering}
    \item Moving a node down the proof to resolve it only once.
  \end{subpart}
  \begin{subpart}{Lowering Units}
    \item Units can always be lowered.
    \item Reduces horizontal irregularities.
  \end{subpart}
  \begin{subpart}{Two traversals}
    \item Collect units with more than one child.
    \item Delete units and fix the proof then reintroduce units.
  \end{subpart}
\end{frame}

\begin{frame}{Sequential Composition}
\centering
\charttitle{\compC}
\include{composition-chart}
\end{frame}


\section{Difficulties}
\subsection{to find a combined algorithm always at least as good as both compositions}

\begin{frame}{Irregular Units}
  \begin{center}
\begin{tikzpicture}[node distance=1.3cm]
  \proofnode[color=addcolor]{luroot}{\only<4>{$\bot$}};

  \proofnode[above left of=luroot]{root}{\alt<3-4>{$\dual{a}$}{$\bot$}};
  \withchildren{root} {r2}{\alt<3-4>{$\dual{a} c$}{$c$}} {a4}{$\dual{c}$};
  \withchildren{r2} {r0}{\alt<8>{$b c$}{$b$}} {r1}{$\dual{b} c$};

  \addchildren{r0} {a0}{$\dual{a} b$} {iu}{$a$};
  \draw [proof edge] (r0) -- (a0);
  \only<1,5->{\draw [proof edge] (r0) -- (iu);}
  \only<2>{\draw [deleted edge] (r0) -- (iu);}

  \addchildren{iu} {a1}{$a c$} {a2}{$a \dual{c}$};
  \draw [proof edge] (iu) -- (a1);
  \only<1-6>{\draw [proof edge] (iu) -- (a2);}
  \only<7>{\draw [deleted edge] (iu) -- (a2);}

  \proofnode[above right of=r1]{a3}{$\dual{a} \dual{b} c$};
  \draw [proof edge] (r1) -- (a3);
  \only<1,5->{\draw [proof edge] (r1) -- (iu);}
  \only<2>{\draw [deleted edge] (r1) -- (iu);}

  \crossnode<3-4>{r0}
  \crossnode<3-4>{r1}
  \only<4>{
    \draw [proof edge,color=addcolor] (luroot) -- (root);
    \draw [proof edge,color=addcolor] (luroot) .. controls ++(1.2,1.5) .. (iu);
  }

  \pivot<6-7>{iu}{$c$}
  \pivot<6->{root}{$c$}
  \crossnode<8>{a2}
  \crossnode<8>{iu}

  \draw (luroot) ++(1.2,-0.5) node {};
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Root and Units Safe Literals}
\begin{subpart}{In RPI.LU, after LU :}
  \item the proof is of the form $\eta \odot_{a_0} \eta_0 \odot_{a_1} \cdots \odot_{a_n} \eta_n$ ;
  \item $\set{\dual{a}_i}{i \leq n}$ is the safe literals for $\eta$ : \emph{the root's safe literals} ;
  \item $\forall i < n$, $\set{\dual{a}_j}{i < j \leq n}$ is the safe literals for $\eta_i$ : \emph{a unit's safe literals}.
\end{subpart}
\begin{overlayarea}{\textwidth}{2em}
  \singleline<2->{\asGoodRPILU}
\end{overlayarea}
\end{frame}

\begin{frame}{Units introduced by RPI}
\begin{center}
\vspace{-3em}
\begin{tikzpicture}[node distance=1.3cm,scale=0.93]
  \rootnode;
  \addchildren{root} {n0}{\alt<3->{$\bot$}{$d$}} {a0}{$\dual{d}$};
  \draw<-2> [proof edge] (root) -- (n0);
  \draw     [proof edge] (root) -- (a0);
  \withchildren{n0} {n1}{\alt<3->{$e$}{$d e$}} {n2}{\alt<3->{$\dual{e}$}{$d \dual{e}$}};

  \withchildren{n1} {a1}{$\dual{b} e$} {u0}{\alt<3->{\alert<3>{$b$}}{$d b$}};
  \proofnode[above right of=n2]{a2}{$\dual{b} \dual{e}$};
  \drawchildren{n2} {u0} {a2};

  \withchildren{u0} {n3}{$a b d$} {a3}{$\dual{a} b$};
  \addchildren{n3} {n4}{$b d c$} {a4}{$\dual{c} a$};
  \draw    [proof edge]   (n3) -- (n4);
  \draw<-2>[proof edge]   (n3) -- (a4);
  \addchildren{n4} {a5}{$a b$}   {a6}{$\dual{a} d c$};
  \draw    [proof edge]   (n4) -- (a5);
  \draw<1> [proof edge]   (n4) -- (a6);
  \draw<2> [deleted edge] (n4) -- (a6);

  \pivot<2>{u0} {$a$}
  \pivot<2>{n4} {$a$}
  \crossnode<3->{a6}

  \crossnode<3->{n4}
  \crossnode<3->{n3}
  \crossnode<3->{a4}
  \crossnode<3->{root}
  \crossnode<3->{a0}
\end{tikzpicture}
\vspace{-2.2em}
\end{center}
\begin{overlayarea}{\textwidth}{1em}
  \singleline<4>{\asGoodLURPI}
\end{overlayarea}
\end{frame}


\section{Combined Equivalents}
\subsection{to both sequential compositions}

\begin{frame}{RPI[3]LU}
\singleline{\asGoodRPILU}
\begin{subpart}{Three traversals}
  \item collect units and compute root and units safe literals ;
  \item compute safe literals and mark edges to be deleted ;
  \item fix the proof and reintroduce units.
\end{subpart}
\end{frame}

\begin{frame}{RPI[3]LU vs RPI.LU}
\centering
\only<+>{\charttitle{\compC} \include{3pass-compress}}
\only<+>{\charttitle{\timeC} \include{3pass-time}}
\end{frame}

\begin{frame}{LUnivRPI}
\begin{subpart}{\asGoodLURPI}
  \item Collect units during fixing (top-down traversal).
\end{subpart}
\addtocounter{beamerpauses}{1}
\begin{subpart}{Problems}<+->
  \item<.-> If a unit $\{a\}$ depends on a unit $\{b\}$ it'll be seen as $\left\{a, \dual{b}\right\}$.
  \item<+-> What to do if we find $\left\{a, \dual{b}\right\}$ after $\{b\}$ ?
  \item<+-> A new algorithm extending LU is needed.
\end{subpart}
\end{frame}

\section{LowerUnivalents}
\subsection{A new algorithm extending LU}

\begin{frame}{Lowering a node}
  \begin{subpart}{The (generalized) problem}
    \item Given $\psi[\eta] \odot_{a_0} \eta_0 \odot_{a_1} \cdots \odot_{a_n} \eta_n$
    \item is $\psi_{\text{\tiny fixed}}[] \odot_a \eta \odot_{a_0} \eta_0 \odot_{a_1} \cdots \odot_{a_n} \eta_n$ equivalent ?
  \end{subpart}
  \begin{subpart}{It depends on}
    \item $\eta$ and its children in $\psi[\eta]$ ;
    \item $\set{a_i}{i \leq n}$ called \emph{the clause of lowered pivots}.
  \end{subpart}
\end{frame}

\begin{frame}{Univalent clause}
  \begin{definition}[Valent literal]
  A literal $a$ is said to be valent w.r.t. a clause $\Delta$ iff $\dual{a}$
  does not occur in $\Delta$.
  \end{definition}

  \begin{definition}[Univalent clause]
  A clause $\Gamma$ is said to be univalent w.r.t. another clause $\Delta$ iff
  there is exactly one literal $a$ in $\Gamma$ which is valent w.r.t.
  $\Delta$.
  \end{definition}

  \addtocounter{beamerpauses}{1}
  \begin{subpart}{A univalent clause isn't sufficient for a node to be lowered}<+->
    \item<+-> $\Delta = \left\{ \dual{a} \right\}$ and $\eta$'s clause $= \left\{ a, b \right\}$
    \item<+-> $\eta$'s child with pivot $a$
    \item<+-> \alert{$\dual{a}$ propagated down the proof}
  \end{subpart}
\end{frame}

\begin{frame}{Active literals}
\begin{definition}[Active literal]
Let's consider a node $\eta$ with clause $\Gamma$. A literal $a$ from
$\Gamma$ is said to be an active literal of $\eta$ iff $a$ is the pivot of
one of $\eta$'s child.
\end{definition}

\begin{theorem}
If all the active literals of a node $\eta$ with clause $\Gamma$ univalent
w.r.t. the clause $\Delta$ of lowered literals either are the valent
literal $a$ or belong to $\Delta$, then $\eta$ can be lowered.
\end{theorem}
\end{frame}

\begin{frame}{Partial regularization}
\begin{subpart}{Deletable node \uncover<3->{(implemented)}}
  \item Suppose all active literals of a node $\eta$ belong to $\Delta$.
  \item Deleting $\eta$ will introduce only duals of literals from $\Delta$.
\end{subpart}
\begin{subpart}<2->{Partial regularization \uncover<3->{(not implemented)}}
  \item If $\Delta = \set{a_i}{i \leq n}$ then $\set{\dual{a}_i}{i \leq n}$ is the root safe literals.
  \item If the pivot of any $\eta$'s child belongs to $\Delta$ then delete the edge.
\end{subpart}
\end{frame}

\begin{frame}{Algorithm}
  \SetKw{aElseIf}{else if}
  \SetKwFor{And}{and}{then}{endif}
  \begin{algorithm}[H]
    $\Delta \leftarrow \varnothing$ \;
    \For{every node $\eta$ in a top-down traversal}{
      Fix $\eta$ \;
      \uIf{every $\eta$'s active literal belongs to $\Delta$}{Delete $\eta$ \;}
    \aElseIf{$\eta$'s clause is univalent w.r.t. $\Delta$ with valent literal $a$}
    \And{$a$ is the only active literal not belonging to $\Delta$}{
        Lower $\eta$ \;
        $\Delta \leftarrow \Delta \cup \{ a \}$ \;
      }
     }
  \end{algorithm}
\end{frame}

\begin{frame}{Comparison with LU}
\centering
\only<+>{\charttitle{\compC} \include{LUniv-compress}}
\only<+>{\charttitle{\axioC} \include{LUniv-axiom}}
\only<+>{\charttitle{\timeC} \include{LUniv-time}}
\end{frame}

\appendix

\begin{frame}{Conclusion}
\charttitle*{Compression ratio comparison}
\renewcommand{\chartscale}{0.85}
\begin{columns}

  \column{0.45\textwidth}
  \centering
  \include{composition-chart}

  \column{0.55\textwidth}
  \centering
  \only<+>{\include{combination-chart}}
  \only<+>{\include{combluniv-chart}}
  
\end{columns}
\end{frame}

\begin{frame}{References}

\begin{subpart}{Skeptik}
\item http://github.com/Paradoxika/Skeptik
\end{subpart}

\begin{subpart}{Bibliography}
\item Fontaine, P., Merz, S., Woltzenlogel Paleo, B.: Compression of
  propositional resolution proofs via partial regularization. In: CADE. Lecture
  Notes in Computer Science, vol. 6803, pp. 237--251. Springer (2011)
\end{subpart}

\end{frame}

\end{document}
